Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-08-18 LaTex/paper.tex
reviewed literature section

The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in stock-out situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from "never-out-of-stock'' items to a service level exceeding e.g. 90\%.

@@

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand  distribution or an integrated approach that seeks to map directly from the available data (historic sales records and further variables) to the ordering decision. This approach is often referred to as "data-driven newsvendor'' and discussed  e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data to derive the operational decisions (the order quantity) directly.  An overview of a range approaches can also be found in \cite{huber2019data}.

Although this approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks. First, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand. This can be used to simulate the performance on a per-item level and e.g. optimize the impact on business strategy decisions on conflicting metrics such as stock out- and waste-rate. By forecasting the full demand distribution as opposed to point estimators such as the expected demand or the median of the distribution, the forecast quality can be evaluated for quantiles of the distribution, including the often extensive tails of the distribution. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e. the mechanism behind the stochastic behavior of the customer demand. However, modelling demand directly is vital if a causal analysis is planned at a later stage or independently, for example to study the effect promotions, advertisements,  price changes or other influencing factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy of the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for. From a practical perspective, separating the demand forecast from the operational decisions (i.e. calculating the  order quantities for the next delivery cycle) also allows to evaluate longer-term planning and reduces the complexity as it avoids coupling the complex delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. This also allows to share long-term demand predictions with other business units or external vendors  and wholesalers to ease their planning for the production and supply-chain processes upstream of the retailer. From the perspective of industrial practice, modelling the demand separately from deriving the subsequent orders  has the additional benefit that multiple retail chains can benefit from any improvement in the model description even if the concrete retailers are unrelated to each other. For example, if a particular effect was identified at some retailer A and included in the machine learning model, the improvement can be rolled out immediately or on request to all other retailers using this system on a planetary scale without having to adapt the underlying machine learning model for each retailer individually.

In contrast to the direct mapping the observed data to ordering decisions, more traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g. \cite{silver1998} for an overview. In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$. In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g. \cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}. Perishable goods are well described by the "newsvendor-problem'' \cite{Edgeworth}, where in the simplest case all stock perishes at the end of the selling period (e.g. a business day). For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly.

The main objective in any of these approaches is to determine the underlying demand distribution. The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g. \cite{alwan2016}) or via sample average approximation (SAA), see e.g. \cite{shapiro2014} for an overview. However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price or advertisements influence, and therefore are highly correlated with, the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e. the  calculation of the optimal order quantity.

However, in order to be able to fully optimize the operational decision, it is critical that one indeed reconstructs a full demand distribution. This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine-learning approaches, will not suffice. Additionally, we need to consider that demand is not i.i.d., but depends on external factors such as season, weather, product life-cycle, and is also actively changed by the retailer by setting a specific price, offering rebates or running advertisements. Additionally, the demand implicitly depends on the location of the retail outlet as well as the specifics of that location, such as product  assortment influencing the choice of possible replacement articles and many more. These complications are the main reason we cannot treat the replenishment process as $n$ independent newsvendor-type problems.

Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc. into account. This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the  full probability distribution or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favourable, as fewer parameters need to be estimated compared to the case of the full distribution. Empirically, one can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modelled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself that follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution for which $\mu = \sigma^2$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

@@

This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer. Typically this means that the full demand distribution has to be estimated per SKU for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a "black-box'' machine learning model, using Cyclic Boosting allows to interpret how each individual prediction was made and what the most important variables are per individual prediction.

Additionally, we show how the cumulative distribution function (CDF) can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution. This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as  order quantities as well as strategic business decisions by the retailer or gain further insights into customer behavior using for example causal modelling.

>>>>

The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from "never-out-of-stock'' items to a service level exceeding e.g. 90\%.

@@

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand distribution or an integrated approach that seeks to map directly from the observed data to the ordering decision. More traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g. \cite{silver1998} for an overview. In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$. In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g. \cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}. Perishable goods are well described by the "newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g. a business day). For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly. The direct approach is often referred to as "data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e. the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of PDFs can be used to simulate the performance on an individual level, and e.g. optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e. lost sales) and waste rate. From a practical perspective, separating the demand forecast from the operational decisions (i.e. calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of industrial practice, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the concrete retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e. the mechanism behind the stochastic behavior of the customer demand, which is vital if a causal analysis is planned, for example a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.

So, the main objective in the solution of the inventory control problem is to determine the underlying demand distribution. The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g. \cite{alwan2016}) or via sample average approximation (SAA), see e.g. \cite{shapiro2014} for an overview. However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price or advertisements influence, and therefore are highly correlated with, the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e. the calculation of the optimal order quantity. Often, there are many individual time series, for example millions of item-store combinations to be forecasted daily in retail demand forecasting, and traditional, univariate time series forecasting methods, like autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}, operate on each of these time series separately. Machine learning, on the contrary, optimizes all individual time series together, improving the forecast quality of the individual time series by exploiting commonalities between them and therefore reducing the variance of the model. This can be interpreted as kind of hierarchical learning, one example of a hierarchical level being the sum of product sales over all locations in the retail demand forecasting use case. And this reduction of variance does not come with an increase of bias, as usual with the bias-variance tradeoff, because we can still take into account the effects on the finest level. Besides the reduction in variance, this modeling on various hierarchical levels can also be interpreted as kind of transfer learning, improving the generalizability of the method. Furthermore, machine learning enables a natural way to include many exogenous variables, simply as features, what reduces also the bias of the model.

In order to be able to fully optimize the operational decision, it is critical to reconstruct a full demand distribution. This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine learning approaches, will not suffice. Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc. into account. This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, one can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modelled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\mu = \sigma^2$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

@@

This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer. Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a "black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.

Additionally, we show how the cumulative distribution function (CDF) can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution. This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as order quantities as well as strategic business decisions by the retailer, and to gain further insights into customer behavior using for example causal modeling.
